### 2024-08-09 mythbusters.qmd
20240727 - mixed models
20240808 - fixes

### Specify Package Options {#sec-mythbustersPackageOptions}
```{r}
emm_options(lmerTest.limit = 100000)
emm_options(pbkrtest.limit = 100000)
```

@@

load(file = "./data/nfl_playerContracts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
load(file = "./data/nfl_espnQBR.RData")
```

```{r}
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
load(file = "./data/nfl_espnQBR.RData")
```

@@

# Subset data
nfl_playerContractYears <- nfl_playerContracts_subset %>% 
  dplyr::filter(played_in_contract_year == TRUE) %>% 
  dplyr::filter(position %in% c("QB","RB","WR","TE")) %>% 
  dplyr::select(player, position, team, contractYear) %>% 
  dplyr::mutate(playerMerge = toupper(str_replace_all(player, "[[:punct:]]", ""))) %>% 
  dplyr::rename(season = contractYear) %>% 
  dplyr::mutate(contractYear = 1)

# Merge with weekly and seasonal stats data
player_stats_weekly_offense <- player_stats_weekly_offense %>% 
  dplyr::mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))
#nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
#  mutate(playerMerge = toupper(str_replace_all(player_display_name, "[[:punct:]]", "")))

player_statsContracts_offense_weekly <- dplyr::full_join(
  player_stats_weekly_offense,
  nfl_playerContractYears,
  by = c("playerMerge", "position_group" = "position", "season")
) %>% 
  dplyr::filter(position_group %in% c("QB","RB","WR","TE"))

#player_statsContracts_offense_seasonal <- full_join(
#  player_stats_seasonal_offense,
#  nfl_playerContractYears,
#  by = c("playerMerge", "position_group" = "position", "season")
#) %>% 
#  filter(position_group %in% c("QB","RB","WR","TE"))

@@

player_statsContracts_offense_weekly <- player_statsContracts_offense_weekly %>% 
  dplyr::arrange(playerMerge, season, season_type, week)

#player_statsContracts_offense_seasonal <- player_statsContracts_offense_seasonal %>% 
#  arrange(playerMerge, season)

@@

# Merge with QBR data
nfl_espnQBR$playerMerge <- paste(nfl_espnQBR$name_first, nfl_espnQBR$name_last, sep = " ") %>% 
  stringr::str_replace_all(., "[[:punct:]]", "") %>% 
  toupper()

nfl_contractYearQBR_weekly <- nfl_playerContractYears %>% 
  dplyr::filter(position == "QB") %>% 
  dplyr::full_join(
    .,
    nfl_espnQBR,
    by = c("playerMerge","team","season")
  )

@@

nfl_contractYearQBR_weekly <- nfl_contractYearQBR_weekly %>% 
  dplyr::arrange(playerMerge, season, season_type, game_week)

nfl_contractYearQBRsubset_weekly <- nfl_contractYearQBR_weekly %>% 
  dplyr::filter(
    season_type == "Regular",
    game_week != "Season Total") %>% 
  dplyr::mutate(game_week = as.numeric(game_week)) %>% 
  dplyr::arrange(playerMerge, season, season_type, game_week)

# Merge with age and experience
nfl_contractYearQBRsubset_weekly <- player_statsContractsSubset_offense_weekly %>% 
  dplyr::filter(position == "QB") %>% 
  dplyr::select(playerMerge, season, week, age, years_of_experience) %>% 
  full_join(
    nfl_contractYearQBRsubset_weekly,
    by = c("playerMerge","season", c("week" = "game_week"))
  ) %>% 
  arrange(player_id, season, week)

@@

# Calculate Per Game Totals
nfl_contractYearQBR_seasonal <- nfl_contractYearQBRsubset_weekly %>% 
  dplyr::group_by(playerMerge, season) %>% 
  dplyr::summarise(
    age = min(age, na.rm = TRUE),
    years_of_experience = min(years_of_experience, na.rm = TRUE),
    qbr = mean(qbr_total, na.rm = TRUE),
    pts_added = mean(pts_added, na.rm = TRUE),
    epa_pass = mean(pass, na.rm = TRUE),
    qb_plays = sum(qb_plays, na.rm = TRUE), # season total
    contractYear = mean(contractYear, na.rm = TRUE),
    games = n(),
    .groups = "drop_last"
  ) %>% 
  dplyr::mutate(
    contractYear = factor(
      contractYear,
      levels = c(0, 1),
      labels = c("no", "yes")
    ))

@@

nfl_contractYearQBR_seasonal <- nfl_contractYearQBR_seasonal %>% 
  group_by(playerMerge) %>%
  mutate(player_id = as.factor(as.character(cur_group_id())))

>>>>

### Specify Package Options {#sec-mythbustersPackageOptions}

```{r}
emm_options(lmerTest.limit = 100000)
emm_options(pbkrtest.limit = 100000)
```

@@

load(file = "./data/nfl_playerContracts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
load(file = "./data/nfl_espnQBR_seasonal.RData")
```

```{r}
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
```

@@

# Subset data
nfl_playerContractYears <- nfl_playerContracts_subset %>% 
  dplyr::filter(played_in_contract_year == TRUE) %>% 
  dplyr::filter(position %in% c("QB","RB","WR","TE")) %>% 
  dplyr::select(player, position, team, contractYear) %>% 
  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player)) %>% 
  dplyr::rename(season = contractYear) %>% 
  dplyr::mutate(contractYear = 1)

# Merge with weekly and seasonal stats data
player_stats_weekly_offense <- player_stats_weekly_offense %>% 
  dplyr::mutate(nameMerge = petersenlab::cleanUpNames(player_display_name, ))
#nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
#  mutate(nameMerge = petersenlab::cleanUpNames(player_display_name, ))

player_statsContracts_offense_weekly <- dplyr::full_join(
  player_stats_weekly_offense,
  nfl_playerContractYears,
  by = c("nameMerge", "position_group" = "position", "season")
) %>% 
  dplyr::filter(position_group %in% c("QB","RB","WR","TE"))

#player_statsContracts_offense_seasonal <- full_join(
#  player_stats_seasonal_offense,
#  nfl_playerContractYears,
#  by = c("nameMerge", "position_group" = "position", "season")
#) %>% 
#  filter(position_group %in% c("QB","RB","WR","TE"))

@@

player_statsContracts_offense_weekly <- player_statsContracts_offense_weekly %>% 
  dplyr::arrange(nameMerge, season, season_type, week)

#player_statsContracts_offense_seasonal <- player_statsContracts_offense_seasonal %>% 
#  arrange(nameMerge, season)

@@

# Merge with QBR data
nfl_espnQBR_seasonal$nameMerge <- paste(nfl_espnQBR_seasonal$name_first, nfl_espnQBR_seasonal$name_last, sep = " ") %>% 
  petersenlab::cleanUpNames(.)

nfl_contractYearQBR_weekly <- nfl_playerContractYears %>% 
  dplyr::filter(position == "QB") %>% 
  dplyr::full_join(
    .,
    nfl_espnQBR_seasonal,
    by = c("nameMerge","team","season")
  )

@@

nfl_contractYearQBR_weekly <- nfl_contractYearQBR_weekly %>% 
  dplyr::arrange(nameMerge, season, season_type, game_week)

nfl_contractYearQBRsubset_weekly <- nfl_contractYearQBR_weekly %>% 
  dplyr::filter(
    season_type == "Regular",
    game_week != "Season Total") %>% 
  dplyr::mutate(game_week = as.numeric(game_week)) %>% 
  dplyr::arrange(nameMerge, season, season_type, game_week)

# Merge with age and experience
nfl_contractYearQBRsubset_weekly <- player_statsContractsSubset_offense_weekly %>% 
  dplyr::filter(position == "QB") %>% 
  dplyr::select(nameMerge, season, week, age, years_of_experience) %>% 
  full_join(
    nfl_contractYearQBRsubset_weekly,
    by = c("nameMerge","season", c("week" = "game_week"))
  ) %>% 
  arrange(player_id, season, week)

@@

# Calculate Per Game Totals
nfl_contractYearQBR_seasonal <- nfl_contractYearQBRsubset_weekly %>% 
  dplyr::group_by(nameMerge, season) %>% 
  dplyr::summarise(
    age = min(age, na.rm = TRUE),
    years_of_experience = min(years_of_experience, na.rm = TRUE),
    qbr = mean(qbr_total, na.rm = TRUE),
    pts_added = mean(pts_added, na.rm = TRUE),
    epa_pass = mean(pass, na.rm = TRUE),
    qb_plays = sum(qb_plays, na.rm = TRUE), # season total
    contractYear = mean(contractYear, na.rm = TRUE),
    games = n(),
    .groups = "drop_last"
  ) %>% 
  dplyr::mutate(
    contractYear = factor(
      contractYear,
      levels = c(0, 1),
      labels = c("no", "yes")
    ))

@@

nfl_contractYearQBR_seasonal <- nfl_contractYearQBR_seasonal %>% 
  group_by(nameMerge) %>%
  mutate(player_id = as.factor(as.character(cur_group_id())))

### 2024-08-09 sports-cognitive-psychology.qmd
20240808 - fixes

```{r}
library("nflreadr")
library("tidyverse")
```

@@

# Downloaded Data
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings.RData")
load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_offense.RData")

# Calculated Data
load(file = "./data/nfl_actualStats_career.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
```

@@

nfl_4thdown %>%
  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2023)) %>%
  ggplot(aes(go_boost, go, color = as.factor(season))) + 
  geom_richtext(data = text_df,   
                aes(x, y, label = label, angle = angle), 
                color = "black", fill = NA, label.color = NA, size = 5) + 
  geom_vline(xintercept = 0) +
  stat_smooth(method = "gam", method.args = list(gamma = 1), formula = y ~ s(x, bs = "cs", k = 10), show.legend = F, se = F, linewidth = 4) +
  # this is just to get the plot to draw the full 0 to 100 range
  geom_hline(yintercept = 100, alpha = 0) +
  geom_hline(yintercept = 0, alpha = 0) +
  ggplot2::theme_classic() +
  labs(x = "Gain in win probability by going for it",
       y = "Go-for-it percentage",
       subtitle = "4th down decisions in 2023 versus 2014, win prob. > 20%",
       title = glue::glue("How <span style='color:red'>math</span> is changing football")) +
  theme(
    legend.position = "none",
    plot.title = element_markdown(size = 22, hjust = 0.5),
    plot.subtitle = element_markdown(size = 14, hjust = 0.5),
    axis.title.x = element_text(size = 14, face="bold"),
    axis.title.y = element_text(size = 14, face="bold")
  ) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(0, 100), expand = c(0,0)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-10, 10), expand = c(0,0)) +
  annotate("text", x= -1.2, y= 70, label = "Should\nkick", color="black", size = 5) +
  annotate("text", x= 1.2, y= 70, label = "Should\ngo for it", color="black", size = 5) +
  geom_segment(
    aes(x = -.1, y = 80, xend = -2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", size = 2
    ) +
  geom_segment(
    aes(x = .1, y = 80, xend = 2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", size = 2
  )
```

>>>>

```{r}
library("nflreadr")
library("tidyverse")
library("ggtext")
```

@@

load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
```

```{r}
load(file = "./data/nfl_pbp.RData")
load(file = "./data/nfl_4thdown.RData")
```

@@

nfl_4thdown %>%
  filter(vegas_wp > .2, between(go_boost, -10, 10), season %in% c(2014, 2023)) %>%
  ggplot(aes(go_boost, go, color = as.factor(season))) + 
  ggtext::geom_richtext(
    data = text_df,   
    aes(x, y, label = label, angle = angle), 
    color = "black",
    fill = NA,
    label.color = NA,
    size = 5) + 
  geom_vline(xintercept = 0) +
  stat_smooth(method = "gam", method.args = list(gamma = 1), formula = y ~ s(x, bs = "cs", k = 10), show.legend = F, se = F, linewidth = 4) +
  # this is just to get the plot to draw the full 0 to 100 range
  geom_hline(yintercept = 100, alpha = 0) +
  geom_hline(yintercept = 0, alpha = 0) +
  ggplot2::theme_classic() +
  labs(x = "Gain in win probability by going for it",
       y = "Go-for-it percentage",
       subtitle = "4th down decisions in 2023 versus 2014, win prob. > 20%",
       title = glue::glue("How <span style='color:red'>math</span> is changing football")) +
  theme(
    legend.position = "none",
    plot.title = element_markdown(size = 22, hjust = 0.5),
    plot.subtitle = element_markdown(size = 14, hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold")
  ) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(0, 100), expand = c(0,0)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-10, 10), expand = c(0,0)) +
  annotate("text", x= -1.2, y= 70, label = "Should\nkick", color="black", size = 5) +
  annotate("text", x= 1.2, y= 70, label = "Should\ngo for it", color="black", size = 5) +
  geom_segment(
    aes(x = -.1, y = 80, xend = -2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", linewidth = 2
    ) +
  geom_segment(
    aes(x = .1, y = 80, xend = 2, yend = 80),
    arrow = arrow(length = unit(0.05, "npc")),
    color = "black", linewidth = 2
  )
```
